---
title: 'Task 1: Spatial Data Visualization'
author: "Katheryn Moya"
date: "3/13/2023"
output: 
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    code_folding: show
    collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(broom)

## spatial data packages

library(sf) #simple features
library(tmap)
library(terra)
```

## Task Instructions:


- For this task, read in the spatial data, and create a professionally formatted and prepared HTML (showing all of your code directly or making it available with code-folding) from an .Rmd in which you: 

  - Make an exploratory interactive map in tmap showing the location of oil spill events included in the data.
  
  - Make a finalized static choropleth map in ggplot in which the fill color for each county depends on the count of inland oil spill events by county for the 2008 oil spill data
  

## Overview:

*Data Citation:*

### Data Wrangling
```{r}
# read in oil spill dataframe

oil_spill_df <- read_csv(here("data", "Oil_Spill_Incident_Tracking.csv")) %>% 
  janitor::clean_names()

# Read in CA county outlines

ca_counties_sf <- read_sf(here("data", "ca_counties", "CA_Counties_TIGER2016.shp")) %>%  # a way to read in vector data to create a simple features object
  janitor::clean_names() %>%  # gets rid of capitals
  select(county_name = name, land_area = aland)

```

```{r}

# determine county sf crs
ca_counties_crs <- st_crs(ca_counties_sf) 

ca_counties_crs #3857

#convert oil spill df to sf

oil_spill_sf <- oil_spill_df %>% 
  drop_na(latitude, longitude) %>% 
  st_as_sf(coords = c("x", "y"), crs = ca_counties_crs)
```


## Exploratory Interactive Map of Oil Spill Events in 2008

```{r}
tmap_mode("view")

  tm_shape(ca_counties_sf) +
  tm_borders() +
  tm_shape(oil_spill_sf) +
  tm_dots(col = "specificlo") +
  tm_layout(legend.outside = TRUE) +
  tm_layout(title = "California Oil Spill Events in 2008", title.size = 1.5) +
  tm_minimap()
    
```

## Static Choropleth Map of Inland Oil Spill Counts in ggplot

```{r}
# wrangle oil spill data

oil_spill_subset <- oil_spill_df %>% 
  filter(inlandmari == "Inland")

# spatially join county and oil spill data

ca_spill_sf<- ca_counties_sf %>% 
  st_join(oil_spill_sf)

spill_counts_sf <- ca_spill_sf %>% 
  filter(inlandmari == "Inland") %>% 
  group_by(county_name) %>% 
  summarize(n_spills = n())

## Plot the results

ggplot(data = spill_counts_sf )  + 
  geom_sf(aes(fill = n_spills), 
          color = "grey30", 
          size = 0.1) +
  scale_fill_gradientn(colors = c("lightgrey", "orange", "red"))
```

