---
title: 'Task 1: California Oil Spill Incident Visualization'
author: "Katheryn Moya"
date: "3/13/2023"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    code_folding: show
    collapsed: yes
---

```{r setup, include=FALSE, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(broom)

## spatial data packages

library(sf) #simple features
library(tmap)
library(terra)
```


## Overview:

Using data provided by the Office of Spill Prevention and Response (OSPR) database, reported incidents of oil spills in California during the year 2008 was visualized. First, every 2008 oil spill location was mapped on an interactive map with CA county borders. The second part of the analysis includes a choropleth map was created to visualize which counties had the highest number of inland oil spill incidents in 2008.

*Data Citation:* Lampinen, Mark (2020). Oil Spill Incident Tracking [ds394]. California Department of Fish and Game, Office of Spill Prevention and Response. https://gis.data.ca.gov/datasets/CDFW::oil-spill-incident-tracking-ds394-1/about


```{r}
# read in oil spill dataframe

oil_spill_df <- read_csv(here("data", "Oil_Spill_Incident_Tracking.csv")) %>% 
  janitor::clean_names()

# Read in CA county outlines

ca_counties_sf <- read_sf(here("data", "ca_counties", "CA_Counties_TIGER2016.shp")) %>%  # a way to read in vector data to create a simple features object
  janitor::clean_names() %>%  # gets rid of capitals
  select(county_name = name, land_area = aland)

```

```{r}

# determine county sf crs
ca_counties_crs <- st_crs(ca_counties_sf) 

 # ca_counties_crs #3857

#convert oil spill df to sf

oil_spill_sf <- oil_spill_df %>% 
  drop_na(latitude, longitude) %>% 
  st_as_sf(coords = c("x", "y"), crs = ca_counties_crs)
```


## Exploratory Interactive Map of Oil Spill Events in 2008

```{r}
tmap_mode("view")

  tm_shape(ca_counties_sf) +
  tm_borders() +
  tm_shape(oil_spill_sf) +
  tm_dots(col = "specificlo") +
  tm_layout(legend.outside = TRUE) +
  tm_layout(title = "California Oil Spill Events in 2008", title.size = 1.5) +
  tm_minimap()
    
```


**Figure 1.** Reported oil spill incidents in California in 2008. The user can click on the point to obtain more information on each event.


## Static Choropleth Map of Inland Oil Spill Counts in ggplot

```{r}
# wrangle oil spill data

oil_spill_subset <- oil_spill_df %>% 
  filter(inlandmari == "Inland")

# spatially join county and oil spill data

ca_spill_sf<- ca_counties_sf %>% 
  st_join(oil_spill_sf)

spill_counts_sf <- ca_spill_sf %>% 
  filter(inlandmari == "Inland") %>% 
  group_by(county_name) %>% 
  summarize(n_spills = n())

## Plot the results

ggplot(data = spill_counts_sf )  + 
  geom_sf(aes(fill = n_spills), 
          color = "grey30", 
          size = 0.1) +
  scale_fill_gradientn(colors = c("lightblue1", "darkslategray4", "darkslategrey")) +
  labs(fill = "Number of Inland Oil Spills")
```

**Figure 2.** California oil spills that occurred inland in 2008 by county. Counties with darker blues depict a greater number of spills compared to the rest of the state. 
